---
title: "assignment_08_02_PothineniKalyan"
author: "PothineniKalyan"
date: "2023-05-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data/r4ds/heights.csv to
```{r }
heights_df <- read.csv("data/r4ds/heights.csv")
```

## Fit a linear model
```{r }
earn_lm <- lm(earn ~ ed + race + height + age + sex, data=heights_df)
summary(earn_lm)
```

## Predict
```{r }
predicted_df <- data.frame(
earn = predict(earn_lm, heights_df),
ed=heights_df$ed, race=heights_df$race, height=heights_df$height,
age=heights_df$age, sex=heights_df$sex
)
```

## Compute deviation (i.e. residuals)
```{r }
mean_earn <- mean(heights_df$earn)
```

## Corrected Sum of Squares Total
```{r }
sst <- sum((mean_earn - heights_df$earn)^2)
```

## Corrected Sum of Squares for Model
```{r }
ssm <- sum((mean_earn - predicted_df$earn)^2)
```

## Residuals
```{r }
residuals <- heights_df$earn - predicted_df$earn
```

## Sum of Squares for Error
```{r }
sse <- sum(residuals^2)
```

## R Squared
```{r }
r_squared <- ssm / sst
```

## Number of observations
```{r }
n <- nrow(heights_df)
```

## Number of regression paramaters
```{r }
p <- 8
```

## Corrected Degrees of Freedom for Model
```{r }
dfm <- p - 1
```

## Degrees of Freedom for Error
```{r }
dfe <- n - p
```

## Corrected Degrees of Freedom Total: DFT = n - 1
```{r }
dft <- n - 1
```

## Mean of Squares for Model: MSM = SSM / DFM
```{r }
msm <- ssm / dfm
```

## Mean of Squares for Error: MSE = SSE / DFE
```{r }
mse <- sse / dfe
```

## Mean of Squares Total: MST = SST / DFT
```{r }
mst <- sst / dft
```

## F Statistic
```{r }
f_score <- msm / mse
```

## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
```{r }
adjusted_r_squared <- 1 - ((1 - r_squared) * (n - 1) / (n - p))
```

## Output the results
```{r }
cat("R-Squared: ", r_squared, "\n")
cat("Adjusted R-Squared: ", adjusted_r_squared, "\n")
cat("F-Statistic: ", f_score, "\n")
cat("p-value: ", pf(f_score, dfm, dfe, lower.tail=FALSE), "\n")
```